\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-11-17 background.tex
Updated paper

Table~\ref{tab:1} presents the SystemJ kernel statements used to program
the control flow. The data flow is programmed in Java. A number of
derived statements also exist (e.g., \texttt{sustain}, \texttt{await},
etc) that make programming reactive systems easier. The \textit{Model of
  Computation} (MoC) of a simple SystemJ program is shown in
Figure~\ref{fig:2}.

SystemJ consists of entities called \textit{Clock Domains} (CD)s, each
running at their individual pace, interacting with the environment via
signals and with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs adheres to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

@@

\begin{figure}[t!]
% \centering
\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
  \begin{scriptsize}
\begin{verbatim}
{
int counter = 0;
while(true){
 abort(A){
  while(true) pause; 
  emit O;
  if(counter != 0) { // do java computation leading to WCRT}
  ++counter;
  } // abort end
 } // loop end
} // SystemJ CD end
\end{verbatim}%
  \end{scriptsize}
\end{minipage}%
\end{SubFloat}
\hspace{1.5cm}%
% \begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
% \includegraphics[scale=0.4]{moc}
% \end{SubFloat}%
% \vspace{-10pt}

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time, requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the
\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
the amount of computation required and schedule, $\Delta$ can vary,
hence, in order to satisfy the implicit restriction imposed by the
synchrony hypothesis -- no input event \textit{from the environment} can
be missed -- one needs to calculate the \textit{Worst Case Reaction
  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
between any two consecutive events on any input, else the synchrony
hypothesis is violated. Techniques for calculating the WCRT of
synchronous programs~\cite{boldt07} exist and this analysis is not the
focus of this paper. The opposite of the WCRT is the \textit{Best Case
  Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
respectively. Figure~\ref{fig:2c} shows a continuously running physical
clock (analogous to a clock in digital hardware), the numerical
annotations on the rising edge of the clock mark the logical ticks from
Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e} for the
numbers show the start and end of the logical ticks, respectively. The
end of a logical tick and the start of the next logical tick happen
together. Figure~\ref{fig:2c} shows the mapping of the logical time to
the physical time. For example, the logical tick \texttt{1} starts at
the first rising edge and completes at the second rising edge of the
physical clock. Where as logical tick \texttt{4} starts at the $4^{th}$
rising edge, but finishes at the $6^{th}$ rising edge of the physical
clock. This difference can be accounted for by the java computations
required when signal \texttt{A} is input by the environment (see
Figure~\ref{fig:2a}). This elasticity is inherent (and elegant) to both
Esterel and SystemJ.

>>>>

Table~\ref{tab:1} presents the SystemJ kernel statements used to program
the control flow. The data flow is programmed in the safe subset of
Java~\cite{scj2013} amenable to formal verification and timing analysis,
while avoiding all the pitfalls of programming in 'C' style languages. A
number of derived statements also exist (e.g., \texttt{sustain},
\texttt{await}, etc) that make programming reactive systems easier. The
\textit{Model of Computation} (MoC) of a simple SystemJ program is shown
in Figure~\ref{fig:2}.

SystemJ consists of entities called \textit{Clock Domains} (CD)s, each
running at their individual pace, interacting with the environment via
signals and with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs adhere to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

@@

\begin{figure}[t!]
% \centering
\begin{SubFloat}{\label{fig:2a}Simple SystemJ program}%...verbatim subfigure
\begin{minipage}[b]{0.3\linewidth}% a minipage to control the width...
  \begin{scriptsize}
\begin{verbatim}
{
int counter = 0;
while(true){
 abort(immediate A){
  while(true) pause; 
 } // abort end
  emit O;
  if(counter != 0) { // java computation leading to WCRT}
  ++counter;
 } // loop end
} // SystemJ CD end
\end{verbatim}%
  \end{scriptsize}
\end{minipage}%
\end{SubFloat}
\hspace{1.5cm}%
% \begin{SubFloat}[Black box]{\label{fig:2b}Logical time in SystemJ}%
% \includegraphics[scale=0.4]{moc}
% \end{SubFloat}%
% \vspace{-10pt}

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time, requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the
\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
the amount of computation required and schedule, $\Delta$ can vary,
hence, in order to satisfy the implicit restriction imposed by the
synchrony hypothesis -- no input event \textit{from the environment} can
be missed -- one needs to calculate the \textit{Worst Case Reaction
  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
between any two consecutive events on any input, else the synchrony
hypothesis is violated. Techniques for calculating the WCRT of
synchronous programs~\cite{boldt07} exist and this analysis is not the
focus of this paper. The opposite of the WCRT is the \textit{Best Case
  Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
respectively. Figure~\ref{fig:2c} shows a continuously running physical
clock (analogous to a clock in digital hardware), the numerical
annotations on the rising edge of the clock mark the logical ticks from
Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e} for the
numbers show the start and end of the logical ticks, respectively. The
end of a logical tick and the start of the next logical tick happen
together. Figure~\ref{fig:2c} shows the mapping of the logical time to
the physical time. For example, the logical tick \texttt{1} starts at
the first rising edge and completes at the second rising edge of the
physical clock, whereas logical tick \texttt{4} starts at the $4^{th}$
rising edge, but finishes at the $6^{th}$ rising edge of the physical
clock. This difference can be accounted for by the Java computations
required when signal \texttt{A} is input by the environment (see
Figure~\ref{fig:2a}). This elasticity is inherent (and elegant) to both
Esterel and SystemJ.

### 2013-11-17 conclusion.tex
Updated paper

In this paper we have described a novel way to introduce real-time in
\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
particular SystemJ, and their subset the synchronous languages like
Esterel. The fundamental idea is to convert real-time delay into logical
delays (\texttt{pause} construct) that interact well with rest of the
constructs in these languages, especially, preemption and channel
communication. We do \textit{not} need or use external timers to
introduce these real-time delays, thereby resolving the problems of
interaction between external timers and preemption. Moreover, we also do
not require specific timer resolutions, thereby making our solution
elegant (in the spirit of GALS/synchronous programming) and flexible. We
rewrite the delay construct into reactive kernel constructs, calculate
the \textit{Worst} and \textit{Base} case reaction times and then
determine \textit{delay} \texttt{d}, number of logical ticks, which is
invariant to the \textit{elasticity} of logical time. To our knowledge
we are the first to introduce non-deterministic delays in such languages
and show their usefulness in designing real-world systems.

>>>>

In this paper we have described a novel way to introduce real-time in
\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
particular SystemJ, and their subset the synchronous languages like
Esterel. The fundamental idea is to convert real-time delay into logical
delays (\texttt{pause} construct) that interact well with rest of the
constructs in these languages, especially, preemption and channel
communication. We do \textit{not} need or use external timers to
introduce these real-time delays, instead we depend upon an underlying
platform that is timing analyzable, thereby resolving the problems of
interaction between external timers and preemption. Moreover, we also do
not require specific timer resolutions, thereby making our solution
elegant (in the spirit of GALS/synchronous programming) and flexible. We
rewrite the delay construct into reactive kernel constructs, calculate
the \textit{Worst} and \textit{Best} case reaction times and then
determine \textit{delay} \texttt{d}, number of logical ticks, which is
invariant to the \textit{elasticity} of logical time. To our knowledge
we are the first to introduce precise and imprecise delays in such
languages and show their usefulness in designing real-world systems.

### 2013-11-17 delays.tex
Updated paper

The fundamental observation is that real-time is converted into logical
time via the \texttt{pause} construct. The rewrite basically maps the
physical notion of time back to the elegant logical notion of time. The
rewrite \textit{delays} a certain number of logical ticks, before
proceeding to the next statement. The number of logical ticks
\texttt{``d''} to \textit{delay} (Figure~\ref{fig:3}) is determined by
the compiler statically at compile time. The value of \texttt{d} is
intricately tied with the WCRT and BCRT of the program and hence the
execution platform.

@@

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{delay} construct. We first
divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
gives us the number of individual logical ticks required to delay the CD
(or synchronous program) by the real-time specification. We always
$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
these functions guarantee that the resultant logical ticks result in
real-time delays between the required range $M-N$. Next, a function $F$
maps these calculated values to a set of equidistant integer points
(values) separated by a unit value -- these points represent all the
logical ticks running at the WCRT and the BCRT, respectively that
satisfy the real-time delay requirements. The intersection of these two
sets gives all the logical ticks that satisfy the real-time requirements
invariant of the logical time and its elasticity.

@@

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT and
  satisfy the required real-time delay.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450..1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can choose any value
  for \texttt{d}.
\end{enumerate*}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time delay. We
think this is an elegant solution, because the technique provides
hard-real time guarantees while preserving the essence (elastic logical
tick) of synchronous and GALS programming prescribed by SystemJ and
Esterel style languages. Moreover, this technique considers non
maximal-parallelism, i.e., the delays in logical ticks is calculated
after scheduling of CDs or synchronous programs has been performed. To
our knowledge we are the first to do so.

@@

The \texttt{delay(M..M)} variant is a little more interesting. Like
before we find sets $S_1$ and $S_2$, and find the intersection of the
two sets to get the value of \texttt{d}. It is possible (and often
likely, as suggested by our experiments in
Section~\ref{sec:experimental-results}) that the resultant set $D$ is
empty (also possible in the case of \texttt{delay(M..N)}, but never
possible in case of \texttt{delay(M)}). In such a case, we automatically
\textit{relax} the upper bound of the \texttt{delay} statement.

@@

\begin{algorithm}[t!]
  \begin{minipage}{1.0\linewidth}
    \SetAlgoLined
    \KwData{$S_2$, $D$, WCRT}
    \KwResult{d}
    \If {$D = \emptyset$} {
      let $j_{0}$ be the first element of set $S_2$, s.t., $|S_2|=Q$\;
      \ShowLn let $N \leftarrow WCRT \times j_0$\;
      let $d \leftarrow j_0$\;
    }
    \Return d\;
    \caption{Calculating the minimum relaxation of the upper real-time
      bound}
    \label{alg:2}
  \end{minipage}
\end{algorithm}

This algorithm results in the smallest relaxation required for the
real-time delay to be satisfied. The algorithm takes as input the WCRT
and set $S_2$, recall that set $S_2$ represents the logical ticks
required to satisfy the real-time requirement at the BCRT. We take the
very first value from set $S_2$ and multiply it with the WCRT to get the
relaxation $N$. The first element of set $S_2$ is returned as the
logical tick delay \texttt{d}.

The fundamental observation is that we have to delay for a minimum of
$M$ units of real-time, hence, under-approximation is out of
question. We can still over-approximate, but to reduce the resultant
error, we should over-approximate by least possible value, which is the
lower bound of set $S_2$. Thus, the lower bound is considered to be the
only element shared between the two sets $S_1$ and $S_2$ and
accordingly, the upper real-time bound is relaxed by the multiplication
of WCRT and the first element of set $S_2$.

@@

% Another is a printer-spooler example borrowed from timed
% CSP~\cite{Schneider:1999:CRT:555233}. The spooler and printer need to
% synchronize using channels. The printer might be unable to print
% depending upon the paper tray, similarly, spooler might take sometime
% to send the job depending upon its size. Such real-time constraints
% can be modeled in SystemJ as below:

% \begin{scriptsize}

% \begin{verbatim}
% {await (job); delay (2..10 ms);  //SPOOLER CD
%  send mid(job) delay (1ms);}
% ><
% {receive mid; delay(1..30 ms); //PRINTER CD
%  emit print(#mid);}
% \end{verbatim}
% \end{scriptsize}

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
% \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){{await(A);}||{delay(1..1 ms);exit(T);}}
\end{verbatim}
% \end{scriptsize}

@@

% \begin{scriptsize}
\begin{verbatim}
suspend{S} {emit START_TIMER(10); await (TIMER); 
 emit O1;}
\end{verbatim}
% \end{scriptsize}

As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
well with the external timer. The above program sends a signal to an
external timer and waits for 10 ms to pass by. Consider what happens
when signals \texttt{S} and \texttt{TIMER} occur in the same logical
tick, the \texttt{await} statement is never executed (due to suspend)
and hence, we enter a deadlock. Such problems are completely avoided in
our technique, because we convert the real-time delays into logical
delays (\texttt{pause} constructs), which interact well with preemption.

>>>>

The fundamental observation is that real-time is converted into logical
time via the \texttt{pause} construct. The rewrite basically maps the
physical notion of time back to the elegant logical notion of time. The
rewrite \textit{delays} a certain number of logical ticks, before
proceeding to the next statement. The number of logical ticks \texttt{d}
to \textit{delay} (Figure~\ref{fig:3}) is determined by the compiler
statically at compile time. The value of \texttt{d} is intricately tied
with the WCRT and BCRT of the program and hence the execution platform.

@@

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{delay} construct. We first
divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
gives us the number of individual logical ticks required to delay the CD
(or synchronous program) by the real-time specification. We always
$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
these functions guarantee that the resultant logical ticks result in
real-time delays between the required range $[M,N]$. Next, a function
$F$ maps these calculated values to a set of equidistant integer points
(values) separated by a unit value -- these points represent all the
logical ticks running at the WCRT and the BCRT, respectively that
satisfy the real-time delay requirements. The intersection of these two
sets gives all the logical ticks that satisfy the real-time requirements
invariant of the logical time and its elasticity.

@@

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT and
  satisfy the required real-time delay.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450,1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can choose any value
  for \texttt{d}.
\end{enumerate*}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time delay. We
think this is an elegant solution, because the technique provides
real-time guarantees while preserving the essence (elastic logical tick)
of synchronous and GALS programming prescribed by SystemJ and Esterel
style languages. Moreover, this technique considers non
maximal-parallelism, i.e., the delay in logical ticks is calculated
after scheduling of CDs or synchronous programs has been performed. To
our knowledge we are the first to do so.

@@

The \texttt{delay(M..M)} variant is a little more interesting. Like
before we find sets $S_1$ and $S_2$, and find the intersection of the
two sets to get the value of \texttt{d}. It is possible (and often
likely, as suggested by our experiments in
Section~\ref{sec:experimental-results}) that the resultant set $D$ is
empty (also possible in the case of \texttt{delay(M..N)}, but never
possible in case of \texttt{delay(M)}). Note that $S_1$ and $S_2$ can
never be empty sets and will always contain at minimum a single
element. In such a case, we automatically \textit{relax} the upper bound
of the \texttt{delay} statement.

@@

\begin{algorithm}[t!]
  \begin{minipage}{1.0\linewidth}
    \SetAlgoLined
    \KwData{$S_2$, $D$, WCRT}
    \KwResult{d}
    \If {$D = \emptyset$} {
      let $j_{0}$ be the first element of set $S_2$\;
      \ShowLn let $N \leftarrow WCRT \times j_0$\;
      let $d \leftarrow j_0$\;
    }
    \Return d\;
    \caption{Calculating the minimum relaxation of the upper real-time
      bound}
    \label{alg:2}
  \end{minipage}
\end{algorithm}

This algorithm results in the \underline{smallest} relaxation required
for the real-time delay to be satisfied. The algorithm takes as input
the WCRT and set $S_2$, recall that set $S_2$ represents the logical
ticks required to satisfy the real-time requirement at the BCRT. We take
the very first value from set $S_2$ and multiply it with the WCRT to get
the relaxation $N$. The first element of set $S_2$ is returned as the
logical tick delay \texttt{d}. The fundamental observation is that we
have to delay for a minimum of $M$ units of real-time, hence,
under-approximation is out of question. We can still over-approximate,
but to reduce the resultant error, we should over-approximate by least
possible value, which is the lower bound of set $S_2$. Thus, the lower
bound is considered to be the only element shared between the two sets
$S_1$ and $S_2$ and accordingly, the upper real-time bound is relaxed by
the multiplication of WCRT and the first element of set $S_2$.

For example, assume $BCRT=5$ and $WCRT=100$, with a required delay of
$(200..200)$. Algorithm~\ref{alg:1}, results in $D$ being an empty set,
since $S_1=[2]$ and $S_2=[40]$, respectively. In such a case, the
resultant relaxation is: $40 \times 100$ using Algorithm~\ref{alg:2},
which results in a time of $4000$ units. Such large over-approximations
can be avoided by breaking the critical paths (worst case ticks) in the
program (by inserting pause constructs), just like in hardware design,
which is very \textit{unsurprising} considering Esterel has its origins
in hardware.

\subsection{The compiler flow}
\label{sec:systemj-tool-chain}

\begin{figure}[t!]
  \centering
  \includegraphics[scale=0.5]{./tool}
  \caption{The tool-chain flow of SystemJ compiler with delays}
  \label{fig:4}
\end{figure}

The interaction of the low-level execution platform and the described
high-level technique of introducing real-time delays in the SystemJ
language is shown in Figure~\ref{fig:4}. We use static-analysis
techniques to find out the WCRT/BCRT of the programs automatically after
rewrites and standard compiler optimizations. We require timing
analyzable platforms, in our case \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}. Moreover, since algorithms~\ref{alg:1}
and~\ref{alg:2} are performed after scheduling for different constraints
such as faster execution time, or reduced power consumption, we,
inherently satisfy the criteria of exploiting non-maximal
parallelism. The low-level analysis guarantees a precise WCRT/BCRT for
the logical tick, which is then used for calculating $d$ in
Figure~\ref{fig:3}.

@@

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
% \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){{await(A);}||
        {delay(1..1 ms);exit(T);}}
\end{verbatim}
% \end{scriptsize}

@@

% \begin{scriptsize}
\begin{verbatim}
suspend(S) {emit START_TIMER(10); 
            await (TIMER);emit O1;}
\end{verbatim}
% \end{scriptsize}

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The above program sends a signal to an external
timer and waits for 10 ms to pass by. Consider what happens when signals
\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
\texttt{await} statement is never executed (due to suspend) and hence,
we enter a deadlock. Such problems are completely avoided in our
technique, because we convert the real-time delays into logical delays
(\texttt{pause} constructs), which interact well with preemption.
